% ============================================================
%  S1 Software Report — Quantitative Trading Platform
%  Compile with: pdflatex S1_Software_Report.tex  (twice)
% ============================================================
\documentclass[12pt, a4paper]{article}

% ── Packages ─────────────────────────────────────────────────────────────────
\usepackage[a4paper, margin=2.5cm]{geometry}
\usepackage{amsmath, amssymb, amsthm}
\usepackage{booktabs}
\usepackage{array}
\usepackage{longtable}
\usepackage{graphicx}
\usepackage{xcolor}
\usepackage{listings}
\usepackage{enumitem}
\usepackage{hyperref}
\usepackage{fancyhdr}
\usepackage{titlesec}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{microtype}
\usepackage{parskip}
\usepackage{setspace}
\usepackage{multirow}
\usepackage{tabularx}
\usepackage{float}

% ── Colours ───────────────────────────────────────────────────────────────────
\definecolor{deepblue}{RGB}{0, 51, 102}
\definecolor{codegreen}{RGB}{0, 128, 0}
\definecolor{codegray}{RGB}{128, 128, 128}
\definecolor{codered}{RGB}{180, 0, 0}
\definecolor{lightgray}{RGB}{245, 245, 245}

% ── Hyperref ──────────────────────────────────────────────────────────────────
\hypersetup{
    colorlinks  = true,
    linkcolor   = deepblue,
    citecolor   = deepblue,
    urlcolor    = deepblue,
    pdftitle    = {S1 Software Report — Quantitative Trading Platform},
    pdfauthor   = {},
}

% ── Code listings ─────────────────────────────────────────────────────────────
\lstset{
    backgroundcolor = \color{lightgray},
    basicstyle      = \ttfamily\small,
    keywordstyle    = \color{deepblue}\bfseries,
    commentstyle    = \color{codegray}\itshape,
    stringstyle     = \color{codegreen},
    breaklines      = true,
    frame           = single,
    framerule       = 0.4pt,
    rulecolor       = \color{codegray},
    numbers         = left,
    numberstyle     = \tiny\color{codegray},
    numbersep       = 8pt,
    showstringspaces= false,
    tabsize         = 4,
    captionpos      = b,
}

% ── Section formatting ────────────────────────────────────────────────────────
\titleformat{\section}{\large\bfseries\color{deepblue}}{}{0em}
    {\thesection\quad}[\titlerule]
\titleformat{\subsection}{\normalsize\bfseries\color{deepblue}}{}{0em}
    {\thesubsection\quad}
\titleformat{\subsubsection}{\normalsize\bfseries}{}{0em}
    {\thesubsubsection\quad}

% ── Header / footer ───────────────────────────────────────────────────────────
\pagestyle{fancy}
\fancyhf{}
\fancyhead[L]{\small\textcolor{deepblue}{S1 Software Report — Quantitative Trading Platform}}
\fancyhead[R]{\small\textcolor{deepblue}{2026}}
\fancyfoot[C]{\thepage}
\renewcommand{\headrulewidth}{0.4pt}

% ── Theorem environments ─────────────────────────────────────────────────────
\theoremstyle{definition}
\newtheorem{defn}{Definition}[section]
\newtheorem{rule_}{Frozen Rule}[section]

% ── Convenience commands ─────────────────────────────────────────────────────
\newcommand{\bp}{\,\text{bp}}
\newcommand{\DVOL}{\textsc{dvol}}
\newcommand{\Nz}{N_{3z}}
\newcommand{\todo}[1]{\textcolor{codered}{[\textbf{TODO:} #1]}}

% ─────────────────────────────────────────────────────────────────────────────
\begin{document}
\onehalfspacing

% ── Title page ───────────────────────────────────────────────────────────────
\begin{titlepage}
\centering
\vspace*{2cm}

{\Huge\bfseries\color{deepblue}
    Quantitative Trading Platform\\[0.4em]
    \large Software Report (S1)
}

\vspace{1.5cm}
\rule{0.6\linewidth}{0.5pt}

\vspace{1cm}
{\large
\begin{tabular}{ll}
    \textbf{Document type}  & Software Report (S1) \\[4pt]
    \textbf{Status}         & Final Draft \\[4pt]
    \textbf{Date}           & May 2026 \\[4pt]
    \textbf{Asset}          & BTCUSDT Perpetual Future (Binance FAPI) \\[4pt]
    \textbf{Signal data}    & Deribit \DVOL{} Index \\[4pt]
    \textbf{Horizon}        & 24-hour hold, daily entry \\[4pt]
    \textbf{Live since}     & 2026-05-14 (paper capital) \\
\end{tabular}
}

\vspace{1.5cm}
\rule{0.6\linewidth}{0.5pt}
\vfill

{\small
This document is the authoritative technical reference for the platform.
It covers system architecture, strategy specifications, cost model,
risk controls, testing methodology, and operational procedures.
}
\end{titlepage}

% ── Table of contents ─────────────────────────────────────────────────────────
\tableofcontents
\newpage

% =============================================================================
\section{Executive Summary}
% =============================================================================

This report describes a research-to-production quantitative trading platform
built around BTCUSDT perpetual futures on Binance. The system is designed
to run multiple independently validated strategies in paper (simulated) trading
mode before any real capital commitment, with a formal lifecycle gate governing
promotion from research to live deployment.

Two strategies have cleared the full validation bar as of the report date:

\begin{table}[H]
\centering
\caption{Strategies in production (paper capital)}
\label{tab:strategy_summary}
\begin{tabular}{llcccc}
\toprule
\textbf{Strategy} & \textbf{Status} & \textbf{OOS Sharpe} & \textbf{Trades} & \textbf{$p$-value} & \textbf{Hold} \\
\midrule
N3\_DVOL\_LONG & Validated & $+2.95$ & 199 & $<0.0005$ & 24h \\
P3\_OIPD\_DD   & Shadow    & $+5.18^*$ & 57   & $0.007$   & 24h \\
\bottomrule
\multicolumn{6}{l}{\small $^*$ Sharpe computed on exclusive trades only (no N3 overlap). Forward review due \(\geq\) 2026-08-14.}
\end{tabular}
\end{table}

A third strategy (DU\_SHORT) was formally killed during validation (OOS
Sharpe $-0.44$, block-bootstrap $p = 1.000$).

The platform consists of a \textbf{FastAPI backend} with a SQLite persistence
layer, a \textbf{React/TypeScript dashboard}, and a \textbf{Docker deployment}
targeting Windows 11. All strategy parameters are frozen in code; any change
requires a new research cycle and a new strategy version.

% =============================================================================
\section{System Architecture}
% =============================================================================

\subsection{Repository Layout}

The repository is structured in layers that map directly to the research
lifecycle:

\begin{lstlisting}[language=bash, caption={Top-level directory structure}]
Quantitative-Framework/
 framework/       Shared cost model, date splits (used by all layers)
 strategies/      Validated/shadow strategy implementations only
 archive/         Retired experimental implementations
 backtest/        Event-driven backtest engine
 live/            Real-time 1m-bar WebSocket simulator
 research/
   active/        Scripts that led to live strategies (N3, P3 trails)
   killed/        Scripts for hypotheses that failed validation + kill log
 data/            Download scripts; raw/ and processed/ parquet files
 results/
   reports/       LaTeX research reports
   n3/            N3 strategy artefacts
   simulation/    Live simulator outputs
   killed/        Artefacts from killed hypotheses
 paper_trading/
   backend/       FastAPI + SQLite backend
   frontend/      React + TypeScript SPA
 tests/           Cross-layer parity + unit test suite
 docs/            Architecture, runbook, this document
\end{lstlisting}

The \texttt{strategies/} directory is a controlled namespace: only strategies
that have cleared the formal OOS validation gate appear there. Experimental
code lives in \texttt{archive/strategies/} and is not imported by any
production path.

\subsection{Component Diagram}

\begin{table}[H]
\centering
\caption{System components and their roles}
\begin{tabularx}{\linewidth}{llX}
\toprule
\textbf{Component} & \textbf{Technology} & \textbf{Role} \\
\midrule
Backend API       & FastAPI + Uvicorn       & Signal evaluation, trade management, risk controls \\
Scheduler         & APScheduler             & 00:00 UTC daily job; 15-min exit check \\
Database          & SQLite + SQLAlchemy     & Immutable trade journal, equity curve, audit log \\
Market data       & Binance FAPI            & Price, funding rate, open interest \\
Volatility data   & Deribit API             & \DVOL{} hourly bars (30d rolling z-score) \\
Frontend          & React + TypeScript      & Dashboard, signals, trades, risk, pipeline \\
Proxy             & Nginx                   & SPA routing; reverse-proxy to backend \\
Container         & Docker Compose          & Two services: \texttt{backend}, \texttt{frontend} \\
CI                & GitHub Actions          & pytest + tsc + build on every push \\
\bottomrule
\end{tabularx}
\end{table}

\subsection{Daily Data Flow}

At 00:00 UTC the scheduler fires the daily signal job:

\begin{enumerate}[leftmargin=2em]
    \item \textbf{Kill switch check} — abort if the system-wide halt is armed.
    \item \textbf{Deribit fetch} — 32 days of hourly \DVOL{} bars; compute 30d
          rolling mean $\mu_D$ and standard deviation $\sigma_D$.
    \item \textbf{Binance fetch} — BTC mark price, funding rate; persist to
          \texttt{market\_data} table.
    \item \textbf{Signal evaluation} — for each active \texttt{BotConfig} row,
          call the appropriate evaluator.
    \item \textbf{Risk gates} (if signal fires):
          per-strategy check $\rightarrow$ portfolio-level check
          $\rightarrow$ kill switch recheck.
    \item \textbf{Execution} — \texttt{open\_trade()} writes to the
          \texttt{trades} table; alert fired.
\end{enumerate}

Every 15 minutes the exit job closes any open trade whose
\texttt{planned\_exit\_timestamp} has passed, fetches the current price and
accumulated funding settlements, and appends an \texttt{equity\_curve} row.

% =============================================================================
\section{Data Pipeline}
% =============================================================================

\subsection{Sources}

\begin{table}[H]
\centering
\caption{External data sources}
\begin{tabular}{lllr}
\toprule
\textbf{Feed} & \textbf{Provider} & \textbf{Granularity} & \textbf{Lookback} \\
\midrule
BTCUSDT 1m klines    & Binance FAPI       & 1 minute  & 30 days rolling \\
Mark price           & Binance FAPI       & On demand & Spot \\
Funding rate         & Binance FAPI       & 8 hours   & Since entry \\
Open interest        & Binance FAPI       & 1 minute  & 24h delta \\
\DVOL{} index        & Deribit REST       & 1 hour    & 32 days rolling \\
\bottomrule
\end{tabular}
\end{table}

\subsection{Data Quality Controls}

Seven automated checks run on every \texttt{GET /api/system/data-quality}
request and are displayed on the dashboard:

\begin{enumerate}[leftmargin=2em]
    \item \textbf{Market data coverage} — at least one \texttt{market\_data} row per day for the last 7 days.
    \item \textbf{Market data freshness} — most recent row within 26 hours.
    \item \textbf{DVOL missing} — no day with a null \DVOL{} value in the last 7 days.
    \item \textbf{DVOL anomaly} — \DVOL{} $\in [10, 200]$; values outside this range flag a data error.
    \item \textbf{Signal coverage} — at least one signal row per active strategy per day.
    \item \textbf{Duplicate signals} — no strategy fires more than once per day.
    \item \textbf{API errors} — zero \texttt{ERROR}-level system log entries in the last 24 hours.
\end{enumerate}

Staleness threshold: any \DVOL{} or price data older than 120 minutes
(\texttt{DATA\_STALE\_MINUTES}) blocks new trade entries via the per-strategy
risk check.

% =============================================================================
\section{Cost Model}
% =============================================================================

All strategy research and live execution use a single shared cost model
defined in \texttt{framework/costs.py}.

\subsection{Model Definition}

\begin{defn}[Transaction cost]
For a single round-trip trade on the Binance USDT-M perpetual (VIP0 tier),
the total cost fraction is:
\begin{equation}
    C_{\text{RT}} = 2\,(\phi + s)
    \label{eq:round_trip_cost}
\end{equation}
where $\phi$ is the exchange fee and $s$ is the estimated half-spread slippage.
\end{defn}

The platform operates exclusively with maker limit orders:

\begin{table}[H]
\centering
\caption{Maker cost model constants}
\begin{tabular}{lrr}
\toprule
\textbf{Component} & \textbf{Fraction} & \textbf{Basis points} \\
\midrule
Maker fee $\phi$       & 0.0002 & 2 bp \\
Slippage estimate $s$  & 0.0001 & 1 bp \\
One-way cost $\phi + s$ & 0.0003 & 3 bp \\
Round-trip $C_{\text{RT}}$ & 0.0006 & 6 bp \\
\bottomrule
\end{tabular}
\end{table}

\subsection{Parity Guarantee}

The cost model is the single source of truth. \textbf{Any discrepancy between
the research path and the live path is a critical bug.} The test suite enforces
this with five assertions in \texttt{tests/test\_parity.py}:

\begin{enumerate}[leftmargin=2em]
    \item $\texttt{MAKER.one\_way\_cost()} = \texttt{ONE\_WAY\_COST}$ in \texttt{app/config.py}
    \item $\texttt{MAKER.round\_trip\_cost()} = \texttt{MAKER\_RT\_COST}$ in \texttt{app/config.py}
    \item $2 \times \texttt{ONE\_WAY\_COST} = \texttt{MAKER\_RT\_COST}$
    \item \texttt{calculate\_pnl()} matches the canonical research formula
    \item The \texttt{BacktestEngine} applies the same round-trip cost as the paper broker
\end{enumerate}

The CI workflow (\texttt{.github/workflows/ci.yml}) runs all root tests on
every push; a failing parity test blocks the merge.

\subsection{Net PnL Formula}

Let $d \in \{+1, -1\}$ denote trade direction (long/short), $P_0$ the
entry mark price, $P_T$ the exit mark price, and $\{f_i\}$ the 8-hour
funding settlements that occurred during the hold window (positive rate
means longs pay shorts).

The implementation uses the \emph{log return} convention throughout —
consistent with the research backtest path (\texttt{research/active/n3/21\_strategy\_backtest.py}):

\begin{equation}
    r_{\text{gross}} = d \cdot \ln\!\left(\frac{P_T}{P_0}\right)
    \label{eq:gross_return}
\end{equation}

\begin{equation}
    r_{\text{fund}} = d \cdot \left(-\sum_{i} f_i\right)
    \label{eq:funding_return}
\end{equation}

\begin{equation}
    r_{\text{net}} = r_{\text{gross}} + r_{\text{fund}} - C_{\text{RT}}
    \label{eq:net_return}
\end{equation}

\begin{equation}
    r_{\text{net,bp}} = r_{\text{net}} \times 10{,}000
    \label{eq:net_return_bp}
\end{equation}

Note: slippage is embedded in the $C_{\text{RT}}$ constant (via
\texttt{ONE\_WAY\_COST} = fee + slippage) rather than tracked as a
separate line item. The \texttt{slippage} field in the trade record is
always zero; the cost is already captured in \texttt{fees}.
Long/short symmetry is enforced by a dedicated test in
\texttt{test\_parity.py}.

% =============================================================================
\section{Strategy Specifications}
% =============================================================================

\subsection{N3 DVOL Long — Validated}

\subsubsection{Hypothesis}

Deribit's implied volatility index (\DVOL{}) captures the forward-looking
fear of options market participants. When \DVOL{} spikes sharply above its
recent baseline (a ``fear spike''), Bitcoin historically mean-reverts upward
over the subsequent 24 hours as the panic unwinds.

\subsubsection{Signal Construction}

Let $D_t$ denote the \DVOL{} index at hour $t$.
The 30-day rolling z-score is:

\begin{equation}
    \Nz(t) = \frac{D_t - \mu_{30}(t)}{\sigma_{30}(t)}
    \label{eq:n3z}
\end{equation}

where $\mu_{30}(t)$ and $\sigma_{30}(t)$ are the rolling mean and standard
deviation over the prior 720 hourly observations (30 days $\times$ 24 hours).

\begin{rule_}[N3 DVOL Long — frozen 2026-05-14]
\label{rule:n3}
A long entry signal fires at daily close (00:00 UTC) if and only if:
\begin{equation}
    \Nz(t) > 0.75 \quad \text{and} \quad D_t \geq 54
    \label{eq:n3_rule}
\end{equation}
The position is held for exactly 24 hours and closed at the then-current
mark price. The short side (N3z $< -0.75$) is not traded — the DU
regime short was independently killed.
\end{rule_}

\subsubsection{Research Trail}

\begin{table}[H]
\centering
\caption{N3 development sequence}
\begin{tabularx}{\linewidth}{llX}
\toprule
\textbf{Script} & \textbf{Stage} & \textbf{Purpose} \\
\midrule
\texttt{17\_dvol\_deep\_dive.py} & IC screen & Walk-forward IC by half-year; block-bootstrap 95\% CI \\
\texttt{18\_n3\_rigorous\_validation.py} & Correction & Non-overlapping return fix (removes autocorrelation inflation) \\
\texttt{19\_n3\_kill\_attempt.py} & Kill attempt & Designed to falsify — OOS 2024-2026; strategy survived \\
\texttt{20\_n3\_regime\_filter.py} & Threshold & DVOL $\geq 54$ selected over rolling median (more robust) \\
\texttt{21\_strategy\_backtest.py} & Backtest & Position-level P\&L with maker costs and funding \\
\texttt{22\_shadow\_backtest.py} & Variants & N3z threshold 0.50 vs 0.75; DVOL 51 vs 54 comparison \\
\bottomrule
\end{tabularx}
\end{table}

\subsubsection{Out-of-Sample Performance}

\begin{table}[H]
\centering
\caption{N3 DVOL Long — OOS performance by period}
\begin{tabular}{lrrrrrr}
\toprule
\textbf{Period} & \textbf{N} & \textbf{Sharpe} & \textbf{PnL (bp)} & \textbf{Max DD (bp)} & \textbf{Win\%} \\
\midrule
OOS 2024-H1 & 48  & $+2.71$ & $+2{,}847$ & $-612$  & 54.2\% \\
OOS 2024-H2 & 52  & $+3.18$ & $+3{,}104$ & $-531$  & 55.8\% \\
OOS 2025-H1 & 51  & $+2.63$ & $+2{,}789$ & $-841$  & 52.9\% \\
OOS 2025-H2 & 48  & $+2.89$ & $+2{,}488$ & $-598$  & 56.3\% \\
\midrule
\textbf{Full OOS} & \textbf{199} & $\mathbf{+2.95}$ & $\mathbf{+11{,}228}$ & $-1{,}612$ & \textbf{54.8\%} \\
\bottomrule
\multicolumn{6}{l}{\small Block-bootstrap $p < 0.0005$ (2000 samples, block size 21 days, seed 42).}
\end{tabular}
\end{table}

\subsection{P3 OI-Price Divergence (DD Regime) — Shadow}

\subsubsection{Hypothesis}

When price and open interest both decline simultaneously (the ``DD regime''),
existing long holders are capitulating. This forced selling creates a
temporary dislocation. Under high volatility conditions (DVOL $\geq 54$),
the subsequent 24-hour mean-reversion provides a long entry opportunity.

\subsubsection{Regime Classification}

Let $\Delta p_{24}$ and $\Delta \text{OI}_{24}$ denote the 24-hour change in
price and open interest respectively. The four regimes are:

\begin{table}[H]
\centering
\caption{OI-Price Divergence regime taxonomy}
\begin{tabular}{cccl}
\toprule
\textbf{Label} & $\Delta p_{24}$ & $\Delta \text{OI}_{24}$ & \textbf{Interpretation} \\
\midrule
DD & $< 0$ & $< 0$ & Long capitulation — mean-reversion candidate \\
DU & $< 0$ & $> 0$ & New shorts entering — continuation possible \\
UD & $> 0$ & $< 0$ & Longs taking profit — weakening trend \\
UU & $> 0$ & $> 0$ & New longs entering — trend continuation \\
\bottomrule
\end{tabular}
\end{table}

\begin{rule_}[P3 OI-Price Divergence — frozen 2026-05-14]
\label{rule:p3}
A long entry signal fires at daily close if and only if:
\begin{equation}
    \Delta p_{24} < 0 \quad \text{and} \quad \Delta \text{OI}_{24} < 0
    \quad \text{and} \quad D_t \geq 54
    \label{eq:p3_rule}
\end{equation}
Hold period: 24 hours.
\end{rule_}

\subsubsection{Exclusive Trade Analysis}

P3 and N3 frequently fire on the same day. The ``exclusive'' subset
excludes dates where N3 also fires, testing whether P3 carries independent
alpha:

\begin{table}[H]
\centering
\caption{P3 exclusive vs.\ overlap performance (OOS 2024–2026)}
\begin{tabular}{lrrrr}
\toprule
\textbf{Subset} & \textbf{N} & \textbf{Sharpe} & \textbf{PnL (bp)} & \textbf{$p$-value} \\
\midrule
All P3 trades    & 142 & $+3.21$ & $+7{,}482$ & $0.002$ \\
P3 exclusive     & 57  & $+5.18$ & $+4{,}109$ & $0.007$ \\
P3 $\cap$ N3     & 85  & $+2.04$ & $+3{,}373$ & $0.041$ \\
\bottomrule
\end{tabular}
\end{table}

The higher Sharpe in the exclusive subset suggests P3 captures a distinct
mechanism from N3 rather than proxying the same signal. Combined
allocation is deferred to the forward review gate ($\geq$ 2026-08-14).

\subsection{Killed Strategies}

\begin{table}[H]
\centering
\caption{Hypotheses formally killed}
\begin{tabular}{llll}
\toprule
\textbf{Signal} & \textbf{Reason killed} & \textbf{OOS Sharpe} & \textbf{$p$-value} \\
\midrule
DU Short (P25)         & Negative OOS edge            & $-0.44$ & $1.000$ \\
CVD Divergence (S6)    & IC insignificant in bootstrap & $< 0$   & $> 0.10$ \\
VPIN Regime (S6)       & No edge $>$ round-trip cost  & —       & —       \\
Signal A (Funding curve) & |IC| below breakeven        & —       & —       \\
Signal B (Informed flow) & |IC| below breakeven        & —       & —       \\
Signal F (Directional VPIN) & IC decays to 0 by lag 5 & —       & —       \\
H1 Basis dislocation   & Execution infeasible ($<$30s) & —      & —       \\
H1 Extended OOS        & IC negative in 2025 period   & —       & —       \\
\bottomrule
\end{tabular}
\end{table}

Full kill logs with exact verdicts and dates are maintained in
\texttt{research/killed/README.md}.

% =============================================================================
\section{Risk Management}
% =============================================================================

\subsection{Kill Switch}

A system-wide kill switch halts all new trade entries immediately. It is
implemented as a database-persisted flag (component \texttt{kill\_switch} in
the \texttt{system\_logs} table) so it survives process restarts. The in-memory
cache is invalidated on each write.

\begin{table}[H]
\centering
\caption{Kill switch states}
\begin{tabular}{lll}
\toprule
\textbf{State} & \textbf{DB record} & \textbf{Effect} \\
\midrule
Armed    & \texttt{message = "ARMED"}    & No new trades; open trades run to expiry \\
Disarmed & \texttt{message = "DISARMED"} & Normal operation \\
\bottomrule
\end{tabular}
\end{table}

\subsection{Per-Strategy Risk Gates}

Before any trade is opened, three per-strategy checks run in order:

\begin{enumerate}[leftmargin=2em]
    \item \textbf{Price validity} — mark price must be positive and non-null.
    \item \textbf{Position limit} — no open trade already exists for this
          market/strategy pair (1 concurrent position maximum).
    \item \textbf{Data freshness} — \DVOL{} data must not be older than
          \texttt{DATA\_STALE\_MINUTES} (120 min); must have at least
          $30 \times 24 = 720$ hourly bars.
\end{enumerate}

\subsection{Portfolio-Level Risk Gates}

Four portfolio-level checks run after the per-strategy checks:

\begin{table}[H]
\centering
\caption{Portfolio risk limits}
\begin{tabular}{llr}
\toprule
\textbf{Gate} & \textbf{Metric} & \textbf{Limit} \\
\midrule
Total positions    & Open trades across all strategies    & $\leq 3$ \\
Same-market cap   & Open trades in same underlying       & $\leq 2$ \\
Daily loss halt   & Closed PnL on current UTC day        & $\geq -500\bp$ \\
Strategy drawdown & Trailing 20-trade drawdown            & $\geq -20\%$ \\
\bottomrule
\end{tabular}
\end{table}

Any failure raises a \texttt{PortfolioRiskBlocked} exception, fires a
\texttt{risk\_blocked} alert, and logs to the audit trail. No exception
propagates to the caller — the daily job continues to the next strategy.

\subsection{Execution Quality Estimator}

An informational execution quality estimate accompanies every trade opening.
It models three cost layers:

\begin{equation}
    C_{\text{total}} = C_{\text{spread}} + C_{\text{adverse}} + C_{\text{latency}} + \phi_{\text{blended}}
    \label{eq:total_cost}
\end{equation}

\begin{itemize}[leftmargin=2em]
    \item $C_{\text{spread}} = \tfrac{1}{2}\text{spread}(1 + \tfrac{D - 54}{54}) + s_{\text{impact}}$
          — half-spread widening with \DVOL{}
    \item $C_{\text{adverse}} = \frac{D}{100\sqrt{365}} \times 10{,}000 \times 5\%$
          — adverse selection fraction of implied daily range
    \item $C_{\text{latency}} = \Delta t_{\text{ms}} \times 0.006 \times \tfrac{D}{54}$
          — price drift during 40 ms API round-trip
    \item $\phi_{\text{blended}} = p_{\text{maker}} \cdot \phi_{\text{maker}} + (1 - p_{\text{maker}}) \cdot \phi_{\text{taker}}$
          — blended fee based on fill probability
\end{itemize}

Maker fill probability decreases linearly above $D = 54$:

\begin{equation}
    p_{\text{maker}} = \max\!\left(0.30,\ 0.80 - 0.10 \cdot \frac{D - 54}{10}\right)
\end{equation}

% =============================================================================
\section{Paper Trading Backend}
% =============================================================================

\subsection{API Surface}

\begin{table}[H]
\centering
\caption{REST API endpoints}
\begin{tabularx}{\linewidth}{llX}
\toprule
\textbf{Method} & \textbf{Path} & \textbf{Purpose} \\
\midrule
GET & \texttt{/api/dashboard}               & Live snapshot: DVOL, N3z, equity, open positions \\
GET & \texttt{/api/signals/latest}          & Most recent evaluation per strategy \\
GET & \texttt{/api/trades}                  & Full trade journal \\
GET & \texttt{/api/performance}             & Equity curve, Sharpe, yearly breakdown \\
GET & \texttt{/api/replay}                  & N3 historical replay vs.\ reference targets \\
GET & \texttt{/api/forward-log/p3}          & P3 shadow log + independence monitor \\
GET & \texttt{/api/risk/state}              & Current portfolio exposure \\
GET/POST & \texttt{/api/risk/kill-switch}   & Read or set kill switch state \\
GET/POST & \texttt{/api/strategies}         & Strategy pipeline status \\
GET & \texttt{/api/system/data-quality}     & 7-point data quality check \\
GET & \texttt{/api/audit}                   & Immutable chronological decision log \\
GET/POST & \texttt{/api/experiments}        & Research run log \\
GET & \texttt{/api/alerts}                  & Alert inbox \\
\bottomrule
\end{tabularx}
\end{table}

\subsection{Database Schema}

The system uses SQLite with SQLAlchemy ORM. Nine tables cover the full
operational and research lifecycle:

\begin{table}[H]
\centering
\caption{Database schema overview}
\begin{tabularx}{\linewidth}{lX}
\toprule
\textbf{Table} & \textbf{Purpose} \\
\midrule
\texttt{market\_data}     & Daily BTCUSDT price, funding rate, \DVOL{} snapshot \\
\texttt{signals}          & One row per evaluation (entry or no-entry + full reason) \\
\texttt{trades}           & Complete trade journal: entry/exit price, PnL attribution \\
\texttt{equity\_curve}    & Running equity, realised PnL, drawdown per trade close \\
\texttt{system\_logs}     & Structured logs (INFO/WARNING/ERROR per component) \\
\texttt{bot\_configs}     & Active market/strategy pairs driving the daily job \\
\texttt{alerts}           & Trade events, data failures, risk blocks (structured) \\
\texttt{strategy\_status} & Formal lifecycle stage per strategy \\
\texttt{experiment\_runs} & Research run metadata: script, params, metrics, verdict \\
\bottomrule
\end{tabularx}
\end{table}

\subsection{Strategy Lifecycle}

\begin{table}[H]
\centering
\caption{Strategy status transitions}
\begin{tabular}{llp{7cm}}
\toprule
\textbf{Status} & \textbf{Trades?} & \textbf{Meaning} \\
\midrule
\texttt{research}   & No  & Hypothesis under active investigation \\
\texttt{candidate}  & No  & Passed IC screen; entering backtest phase \\
\texttt{shadow}     & Yes & In production with paper capital; 3-month forward log required \\
\texttt{validated}  & Yes & Forward review passed; eligible for real capital \\
\texttt{paused}     & No  & Drawdown or manual pause; awaiting review \\
\texttt{killed}     & No  & Formally retired; code moved to \texttt{research/killed/} \\
\bottomrule
\end{tabular}
\end{table}

% =============================================================================
\section{Testing}
% =============================================================================

\subsection{Test Suite Overview}

\begin{table}[H]
\centering
\caption{Test coverage summary}
\begin{tabular}{llr}
\toprule
\textbf{Module} & \textbf{File} & \textbf{Tests} \\
\midrule
Parity (cross-layer) & \texttt{tests/test\_parity.py}             & 26 \\
N3 signal frozen rule & \texttt{tests/test\_n3\_signal.py}         & 13 \\
P3 signal frozen rule & \texttt{backend/tests/test\_p3\_signal.py} & 19 \\
PnL calculator       & \texttt{tests/test\_pnl.py}                & 15 \\
Paper broker         & \texttt{tests/test\_paper\_broker.py}      & 23 \\
Risk checks          & \texttt{tests/test\_risk.py}               & 14 \\
Integration cycle    & \texttt{backend/tests/test\_integration.py} & 7 \\
Replay certification & \texttt{backend/tests/test\_replay\_certification.py} & 9 \\
\midrule
\textbf{Total} & & \textbf{126} \\
\bottomrule
\end{tabular}
\end{table}

\subsection{Parity Tests}

The parity test suite (\texttt{test\_parity.py}) is the most critical component.
It enforces that the live paper trading system uses exactly the same cost
constants, PnL formula, and signal evaluation logic as the research path.
All 26 assertions must pass before any code can be merged. A failure indicates
the live rule has drifted from the validated research rule and trading must
halt until it is corrected.

\subsection{Integration Tests}

\texttt{test\_integration.py} runs a full signal-to-close cycle against an
in-memory SQLite database with no external API calls:

\begin{enumerate}[leftmargin=2em]
    \item \texttt{open\_trade()} creates a Trade row
    \item Per-strategy risk check blocks a duplicate entry
    \item Portfolio cap blocks entries when the global limit is hit
    \item Kill switch arms, blocks, disarms, and clears correctly
    \item \texttt{close\_trade()} updates all fields and appends an equity curve row
    \item Short-side PnL sign is correct (losing short is negative)
    \item Fees field exactly equals \texttt{MAKER\_RT\_COST} for a flat trade
\end{enumerate}

\subsection{Replay Certification}

\texttt{test\_replay\_certification.py} provides two layers:

\begin{itemize}[leftmargin=2em]
    \item \textbf{Unit layer (always runs)} — 8 of the 9 tests verify that the
          \texttt{\_build\_trades()} and \texttt{\_compute\_stats()} functions
          correctly apply the frozen rule to synthetic DataFrames.
    \item \textbf{Integration layer (conditional)} — skipped automatically when
          local parquet files are absent (CI). When files are present,
          asserts that \texttt{GET /api/replay} matches reference targets within
          a 5\% tolerance:
          $n \approx 199$, Sharpe $\approx +2.95$, PnL $\approx +11{,}228\bp$.
\end{itemize}

% =============================================================================
\section{Deployment}
% =============================================================================

\subsection{Platform}

\begin{table}[H]
\centering
\caption{Deployment environment}
\begin{tabular}{ll}
\toprule
\textbf{Property} & \textbf{Value} \\
\midrule
OS              & Windows 11 Pro (build 26100) \\
Container engine & Docker Desktop with WSL2 backend \\
Backend runtime  & Python 3.11 inside \texttt{python:3.11-slim} \\
Frontend runtime & Node 20 (build); Nginx Alpine (serve) \\
Database         & SQLite (single file, Docker named volume) \\
Network          & Bridge; backend on 8000, frontend on 80 \\
\bottomrule
\end{tabular}
\end{table}

\subsection{Environment Variables}

All optional configuration is supplied via \texttt{paper\_trading/.env}
(template: \texttt{.env.example}). No required variables exist — the system
runs with defaults for local development. Optional email alerts require all
five \texttt{ALERT\_*} variables to be set simultaneously.

\subsection{Continuous Integration}

GitHub Actions runs three jobs on every push and pull request:

\begin{enumerate}[leftmargin=2em]
    \item \textbf{pytest (root)} — parity and unit tests against Python 3.11.
    \item \textbf{pytest (backend)} — signal, integration, and replay tests.
    \item \textbf{tsc + build} — TypeScript type check and React production build.
\end{enumerate}

A merge is blocked if any job fails.

% =============================================================================
\section{Operational Procedures}
% =============================================================================

A full operational runbook is maintained separately at \texttt{docs/runbook.md}.
The key procedures are summarised here for reference.

\subsection{Severity Classification}

\begin{table}[H]
\centering
\caption{Incident severity levels}
\begin{tabular}{lll}
\toprule
\textbf{Severity} & \textbf{Condition} & \textbf{Immediate action} \\
\midrule
S1 — Halt    & Wrong trade direction; logic bypass; DB corrupt & Arm kill switch immediately \\
S2 — Degrade & Data feed down; scheduler missed             & Arm if signal window $<$ 2h \\
S3 — Monitor & Single stale row; email alerts broken        & Investigate at convenience \\
\bottomrule
\end{tabular}
\end{table}

\subsection{P3 Forward Review Gate}

P3\_OIPD\_DD must be reviewed on or after \textbf{2026-08-14}. Promotion to
\texttt{validated} requires all of:

\begin{itemize}[leftmargin=2em]
    \item Exclusive forward trades $\geq 20$
    \item Forward Sharpe $> 1.0$
    \item Block-bootstrap $p \leq 0.05$ on the forward log
    \item All parity tests green
\end{itemize}

Failure on any criterion triggers promotion to \texttt{killed} and movement
of research scripts to \texttt{research/killed/}.

% =============================================================================
\section{Known Limitations and Future Work}
% =============================================================================

\begin{table}[H]
\centering
\caption{Current limitations}
\begin{tabularx}{\linewidth}{lX}
\toprule
\textbf{Area} & \textbf{Limitation} \\
\midrule
Database        & SQLite does not support concurrent writes; not suitable for multi-process deployment \\
Position sizing & Fixed notional (\$1k equivalent); no Kelly or volatility-targeted sizing \\
Short side      & DU regime short killed; no validated short strategy currently \\
Multi-asset     & Single market (BTCUSDT perp) only; no cross-asset correlation controls \\
Execution       & Maker fill probability is a model estimate; not measured from real fills \\
Live capital    & Platform operates on paper capital only; real-capital deployment requires further review \\
\bottomrule
\end{tabularx}
\end{table}

\begin{table}[H]
\centering
\caption{Candidate future work}
\begin{tabularx}{\linewidth}{lX}
\toprule
\textbf{Item} & \textbf{Description} \\
\midrule
P3 validation & Forward review on 2026-08-14; combine with N3 if independent alpha confirmed \\
Sizing model  & Volatility-targeting: scale notional inversely with recent \DVOL{} \\
Short side    & Re-examine DU regime with updated 2026 data; test alternative short triggers \\
PostgreSQL    & Migrate from SQLite to allow concurrent reads from dashboard and scheduler \\
Alert escalation & PagerDuty or similar integration for S1 incidents \\
\bottomrule
\end{tabularx}
\end{table}

% =============================================================================
\section{Glossary}
% =============================================================================

\begin{description}[leftmargin=3cm, style=nextline]
    \item[\DVOL{}]
        Deribit Volatility Index — 30-day forward-looking implied volatility
        for Bitcoin, expressed as an annualised percentage (e.g.\ 54 = 54\%).

    \item[$\Nz$]
        30-day rolling z-score of the hourly \DVOL{} index. A value above 0.75
        indicates \DVOL{} is more than 0.75 standard deviations above its
        recent 30-day mean — a ``fear spike.''

    \item[DD regime]
        OI-Price Divergence regime where both 24h price change and 24h open
        interest change are negative simultaneously.

    \item[Round-trip cost]
        Total transaction cost for entering and exiting one position:
        $C_{\text{RT}} = 6\bp$ (maker path).

    \item[Block bootstrap]
        A resampling method that preserves temporal autocorrelation by
        sampling contiguous blocks rather than individual observations.
        Used to compute $p$-values for trading signal IC tests.

    \item[IC]
        Information Coefficient — Spearman rank correlation between the
        signal value and the subsequent return.

    \item[OOS]
        Out-of-sample — data that was not used during signal development
        or threshold selection.

    \item[Exclusive trades]
        Trades taken by P3 on days when N3 does \emph{not} also fire,
        used to measure P3's independent alpha.

    \item[Kill switch]
        A database-persisted flag that halts all new trade entries
        system-wide, surviving process restarts.

    \item[Parity test]
        An assertion in \texttt{tests/test\_parity.py} that confirms the live
        implementation uses the exact same formula as the research path.
        Enforced on every push via the CI workflow.
\end{description}

% =============================================================================
\appendix
\section{Frozen Rule Quick Reference}
% =============================================================================

\begin{table}[H]
\centering
\caption{Frozen strategy parameters — do not modify without a new research cycle}
\begin{tabular}{llrr}
\toprule
\textbf{Parameter} & \textbf{Symbol} & \textbf{N3} & \textbf{P3} \\
\midrule
N3z entry threshold      & $\tau_{N3z}$  & $0.75$  & — \\
\DVOL{} level threshold  & $\tau_D$       & $54.0$  & $54.0$ \\
\DVOL{} lookback (hours) & $L_D$          & $720$   & — \\
Hold period (hours)      & $H$            & $24$    & $24$ \\
Side                     & —              & Long    & Long \\
Maker round-trip cost    & $C_\text{RT}$  & $6\bp$  & $6\bp$ \\
\bottomrule
\end{tabular}
\end{table}

% =============================================================================
\section{Test Coverage Detail}
% =============================================================================

\begin{table}[H]
\centering
\caption{\texttt{test\_parity.py} — full assertion list}
\begin{tabularx}{\linewidth}{lX}
\toprule
\textbf{Test} & \textbf{Assertion} \\
\midrule
\texttt{test\_one\_way\_cost\_matches}       & $\texttt{MAKER.one\_way\_cost()} = \texttt{ONE\_WAY\_COST}$ \\
\texttt{test\_rt\_cost\_matches}             & $\texttt{MAKER.round\_trip\_cost()} = \texttt{MAKER\_RT\_COST}$ \\
\texttt{test\_rt\_is\_double\_one\_way}      & $\texttt{MAKER\_RT\_COST} = 2 \times \texttt{ONE\_WAY\_COST}$ \\
\texttt{test\_6bp\_sanity}                   & $\texttt{MAKER\_RT\_COST} = 0.0006$ \\
\texttt{test\_pnl\_formula[long,flat,*]}     & $d{=}+1$: log return = 0 $\Rightarrow$ net PnL $= -C_\text{RT}$ (parametrised) \\
\texttt{test\_pnl\_formula[short,flat,*]}    & $d{=}-1$: log return = 0 $\Rightarrow$ net PnL $= -C_\text{RT}$ (parametrised) \\
\texttt{test\_bp\_conversion}               & $r_{\text{net,bp}} = r_{\text{net}} \times 10{,}000$ (Eq.~\ref{eq:net_return_bp}) \\
\texttt{test\_fees\_equal\_rt\_cost}         & \texttt{fees} $= 2\times\texttt{ONE\_WAY\_COST} = \texttt{MAKER\_RT\_COST}$ (flat trade) \\
\texttt{test\_long\_short\_symmetry[*]}      & Long and short PnL are sign-symmetric (parametrised) \\
\texttt{test\_n3\_signal\_decision[*]}       & Evaluator matches direct N3z computation (parametrised) \\
\texttt{test\_n3z\_stored\_in\_signal}       & \texttt{Signal.n3\_z} field is populated \\
\texttt{test\_n3\_side\_is\_long}            & Signal fires long only \\
\texttt{test\_n3\_hold\_is\_24h}             & Hold period is exactly 24 hours \\
\texttt{test\_backtest\_cost}               & \texttt{BacktestEngine} applies same 6bp RT cost \\
\bottomrule
\end{tabularx}
\end{table}

\end{document}
